---
title: "Sixth Blog Post"
subtitle: "Building Map analysis and Thesis"
author: "TEAM NA"
date: "2024-04-22"
draft: false
---

We imported the Newyork Maps shf to create heat maps to analyze where the highest amount of Assault is in NYC, 


Temporary Thesis: the Crime rate of Assault 3 will increase as temperature rises in the New York City.


```{r include=FALSE}
library(dplyr)
library(tidyverse) 
library(ggplot2)
library(dplyr)
library(kableExtra)
library(ggcorrplot)
library(ggmap)
library(leaflet)
library(patchwork)
library(tigris)
library(sp)
library(sf)
library(data.table)
library(lubridate)
library(gridExtra)

ds <- read_rds("dataset/load_and_clean_data.rds")
```



We found that when season changes from summer to winter, the crime density gradually decreases.

```{r echo=FALSE, message=FALSE, warning= FALSE}
ds <- readRDS("dataset/load_and_clean_data.rds")
ds <- as.data.table(ds)
ds$ARREST_DATE <- ymd(ds$ARREST_DATE)
ds$ARREST_month <- month(ds$ARREST_DATE)

ds_sub <- ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES",]
ds_sub_winter <- ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"&ARREST_month %in%c(1,2,12) ,]
ds_sub_summer <- ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"&ARREST_month %in%c(6,7,8) ,]
ds_sub_low <-ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"& temp<14.8 ,]
ds_sub_high <-ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"& temp>14.8 ,]
ds_sub_spring <-ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"& ARREST_month %in%c(3,4,5) ,]
ds_sub_fall <- ds[OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES"&ARREST_month %in%c(9,10,11) ,]


ds_plot <- ds_sub[,.(count = .N),ARREST_PRECINCT ]
ds_plot_winter <- ds_sub_winter[,.(count = .N),ARREST_PRECINCT ]
ds_plot_summer <-ds_sub_summer[,.(count = .N),ARREST_PRECINCT ]
ds_plot_low <- ds_sub_low[,.(count = .N),ARREST_PRECINCT ]
ds_plot_high <- ds_sub_high[,.(count = .N),ARREST_PRECINCT ]
ds_plot_spring <-ds_sub_spring[,.(count = .N),ARREST_PRECINCT ]
ds_plot_fall <-ds_sub_fall[,.(count = .N),ARREST_PRECINCT ]

map_data <- st_read("dataset/geo_export_87553334-fdcd-4d0c-8ab2-43a5e2975733.shp")
map_data <-as(map_data, "Spatial")
map_data_all <- sp::merge(map_data,ds_plot,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_winter <- sp::merge(map_data,ds_plot_winter,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_summer <- sp::merge(map_data,ds_plot_summer,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_low <- sp::merge(map_data,ds_plot_low,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_high <- sp::merge(map_data,ds_plot_high,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_spring <- sp::merge(map_data,ds_plot_spring,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)
map_data_fall <- sp::merge(map_data,ds_plot_fall,by.x = "precinct",by.y ="ARREST_PRECINCT",all.x = T)

map_data_all <- as(map_data_all, "sf")
map_data_winter <- as(map_data_winter, "sf")
map_data_summer <-as(map_data_summer, "sf")
map_data_low <-as(map_data_low, "sf")
map_data_high  <-as(map_data_high , "sf")
map_data_spring <-as(map_data_spring , "sf")
map_data_fall <- as(map_data_fall , "sf")

```


```{r echo=FALSE, message=FALSE, warning= FALSE}
ggplot()+
  geom_sf(data = map_data_all,colour = "gray60",size = 0.1,aes(fill = count) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_distiller(palette = "YlOrRd",direction= 1) +
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("ALL YEAR")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
winter<- ggplot()+
  geom_sf(data = map_data_winter,colour = "gray60",size = 0.1,aes(fill = count) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_distiller(palette = "YlOrRd",direction= 1) +
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("winter")+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning= FALSE}
summer<- ggplot()+
  geom_sf(data = map_data_summer,colour = "gray60",size = 0.1,aes(fill = count) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_distiller(palette = "YlOrRd",direction= 1) +
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("summer")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
ggplot()+
  geom_sf(data = map_data_low,colour = "gray60",size = 0.1,aes(fill = count) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_distiller(palette = "YlOrRd",direction= 1) +
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("low")+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r echo=FALSE, message=FALSE, warning= FALSE}
ggplot()+
  geom_sf(data = map_data_high,colour = "gray60",size = 0.1,aes(fill = count) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_distiller(palette = "YlOrRd",direction= 1) +
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("high")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
map_data_all$count_new<-cut(map_data_all$count
                            ,breaks=c(-Inf,200,400,600,800,1000,Inf)
                            ,labels = c(
                              "[0,200 )"
                              ,"[200,400)"
                              ,"[400,600)"
                              ,"[600,800)"
                              ,"[800,1000)"
                              ,"[1000+)"
                              
                            ))
ggplot()+
  geom_sf(data = map_data_all,colour = "gray60",size = 0.1,aes(fill = count_new) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_brewer(palette = "YlOrRd")+
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("All year")+
  theme(plot.title = element_text(hjust = 0.5))

```

```{r echo=FALSE, message=FALSE, warning= FALSE}
map_data_winter$count_new<-cut(map_data_winter$count
                            ,breaks=c(-Inf,50,100,150,200,250,300,Inf)
                            ,labels = c(
                              "[0,50 )"
                              ,"[50,100)"
                              ,"[100,150)"
                              ,"[150,200)"
                              ,"[200,250)"
                              ,"[250,300)"
                              ,"[300+)"
                              
                              
                            ))

winter <-ggplot()+
  geom_sf(data = map_data_winter,colour = "gray60",size = 0.1,aes(fill = count_new) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_brewer(palette = "YlOrRd")+
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("Winter")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
map_data_summer$count_new<-cut(map_data_summer$count
                            ,breaks=c(-Inf,50,100,150,200,250,300,Inf)
                            ,labels = c(
                              "[0,50 )"
                              ,"[50,100)"
                              ,"[100,150)"
                              ,"[150,200)"
                              ,"[200,250)"
                              ,"[250,300)"
                              ,"[300+)"
                              
                            ))

summer<-ggplot()+
  geom_sf(data = map_data_summer,colour = "gray60",size = 0.1,aes(fill = count_new) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_brewer(palette = "YlOrRd")+
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("Summer")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
map_data_spring$count_new<-cut(map_data_spring$count
                            ,breaks=c(-Inf,50,100,150,200,250,300,Inf)
                            ,labels = c(
                              "[0,50 )"
                              ,"[50,100)"
                              ,"[100,150)"
                              ,"[150,200)"
                              ,"[200,250)"
                              ,"[250,300)"
                              ,"[300+)"
                              
                            ))

spring<-ggplot()+
  geom_sf(data = map_data_spring,colour = "gray60",size = 0.1,aes(fill = count_new) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_brewer(palette = "YlOrRd")+
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("Spring")+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning= FALSE}
map_data_fall$count_new<-cut(map_data_fall$count
                            ,breaks=c(-Inf,50,100,150,200,250,300,Inf)
                            ,labels = c(
                              "[0,50 )"
                              ,"[50,100)"
                              ,"[100,150)"
                              ,"[150,200)"
                              ,"[200,250)"
                              ,"[250,300)"
                              ,"[300+)"
                              
                            ))

fall<-ggplot()+
  geom_sf(data = map_data_fall,colour = "gray60",size = 0.1,aes(fill = count_new) )+
  # scale_fill_continuous(low = "#FF9797", high = "#600000",na.value="white")+
  scale_fill_brewer(palette = "YlOrRd")+
  #scale_fill_manual(palette = "YlOrRd")+
  theme( axis.ticks=element_blank(), 
         axis.title=element_blank(), 
         panel.border = element_blank(), 
         axis.text=element_blank() , 
         panel.background = element_blank()
         
  )+ggtitle("Fall")+
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning= FALSE}
grid.arrange(spring, fall, summer, winter, ncol = 2, nrow = 2)
```

These maps reveal a clear pattern in "Assault 3" incidents: they peak
during the summer months, likely due to the increase in outdoor
activities and social interactions that warmer weather encourages.
Spring and Fall act as transitional periods with moderate crime rates,
suggesting a correlation with the gradual changes in weather and
daylight hours, which influence how much time people spend outside. The
winter months show a marked decrease in crime rates, possibly due to the
cold weather discouraging outdoor activity. These trends indicate that
environmental factors, such as temperature and seasonality, may play a
significant role in the fluctuation of crime rates, necessitating
seasonally adjusted strategies for law enforcement and community safety
programs.