---
title: "Seventh Blog Post"
subtitle: ""
author: "TEAM 14"
date: "2024-04-29"
draft: false
---


```{r include=F}
library(tidyverse)
library(lubridate)
library(patchwork)
data <- read_rds("dataset/load_and_clean_data.rds")
data$ARREST_DATE <- as.Date(data$ARREST_DATE, "%m/%d/%Y")
```

```{r echo=FALSE}
get_season <- function(date) {
  month <- month(date)
  year <- year(date)
  if (month %in% c(12, 1, 2)) {
    season <- "Winter"
  } else if (month %in% c(3, 4, 5)) {
    season <- "Spring"
  } else if (month %in% c(6, 7, 8)) {
    season <- "Summer"
  } else {
    season <- "Autumn"
  }
  paste(season, year)
}
```

```{r echo=F}
#cumulative counts 
data$Season <- sapply(data$ARREST_DATE, get_season)
data_filtered <- filter(data, year(ARREST_DATE) %in% c(2022, 2023))

season_counts <- data_filtered %>%
  group_by(Season) %>%
  summarise(Count = n()) %>%
  mutate(Year = substr(Season, 8, 11)) %>%
  mutate(Season = factor(substr(Season, 1, 6), levels = c("Spring", "Summer", "Autumn","Winter"))) %>%
  arrange(Season)

plot_count<-ggplot(season_counts, aes(x = Season, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("2022" = "steelblue", "2023" = "lightgreen")) +
  theme_minimal() +
  labs(title = "Total Crime Counts per Season", 
       x = "Season", 
       y = "Number of Incidents",
       fill = "Year") 

```

```{r echo=FALSE}
#felcony
data$Season <- sapply(data$ARREST_DATE, get_season)
felony_assault_data <- filter(data, OFNS_DESC == "FELONY ASSAULT", year(ARREST_DATE) %in% c(2022, 2023))

season_felony_count <- felony_assault_data %>%
  group_by(Season) %>%
  summarise(Count = n())%>%
  mutate(Year = substr(Season, 8, 11)) %>%
  mutate(Season = factor(substr(Season, 1, 6), levels = c("Spring", "Summer", "Autumn","Winter"))) %>%
  arrange(Season)
  
plotfa<- ggplot(season_felony_count, aes(x = Season, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("2022" = "steelblue", "2023" = "lightgreen")) +
  theme_minimal() +
  labs(title = "Felony Assault", 
       x = "Season", 
       y = "Number of Incidents",
       fill = "Year")

```

```{r echo=F}
#petit larceny
data$Season <- sapply(data$ARREST_DATE, get_season)
petit_larceny_data <- filter(data, OFNS_DESC == "PETIT LARCENY", year(ARREST_DATE) %in% c(2022, 2023))

petit_larceny_counts <- petit_larceny_data %>%
  group_by(Season) %>%
  summarise(Count = n()) %>%
  mutate(Year = substr(Season, 8, 11)) %>%
  mutate(Season = factor(substr(Season, 1, 6), levels = c("Spring", "Summer", "Autumn", "Winter"))) %>%
  arrange(Season)

plotpl<- ggplot(petit_larceny_counts, aes(x = Season, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("2022" = "steelblue", "2023" = "lightgreen")) +
  theme_minimal() +
  labs(title = "Petit Larceny", 
       x = "Season", 
       y = "Number of Incidents",
       fill = "Year")
```

```{r echo=F}
# assault 3
data$Season <- sapply(data$ARREST_DATE, get_season)
assault_3_data <- filter(data, OFNS_DESC == "ASSAULT 3 & RELATED OFFENSES", year(ARREST_DATE) %in% c(2022, 2023))

assault_3_counts <- assault_3_data %>%
  group_by(Season) %>%
  summarise(Count = n()) %>%
  mutate(Year = substr(Season, 8, 11)) %>%
  mutate(Season = factor(substr(Season, 1, 6), levels = c("Spring", "Summer", "Autumn", "Winter"))) %>%
  arrange(Season)

 plota3<- ggplot(assault_3_counts, aes(x = Season, y = Count, fill = Year)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("2022" = "steelblue", "2023" = "lightgreen")) +
  theme_minimal() +
  labs(title = "Assault 3 ", 
       x = "Season", 
       y = "Number of Incidents",
       fill = "Year")

```
```{r echo=FALSE}
combine_plot_crime<-plot_count+plota3+plotfa+plotpl+
   plot_layout(nrow = 2, ncol = 2)
combine_plot_crime
```

These 4 graphs shows the total crime counts per season and top 3 crime categories per seasons. It all shows a trend that the crime rate will be higher in Summer and Spring, then it will gradually go lower in Autumn and Winter in 2023.This data suggests a correlation between warmer weather and increased crime rates in New York City, indicating that incidents tend to escalate during warmer periods.


```{r echo=FALSE, include=F}
#2022 female vs male 
get_season <- function(date) {
  month <- month(date)
  if (month %in% c(12, 1, 2)) {
    season <- "Winter"
  } else if (month %in% c(3, 4, 5)) {
    season <- "Spring"
  } else if (month %in% c(6, 7, 8)) {
    season <- "Summer"
  } else {
    season <- "Autumn"
  }
  return(season)
}


data$Season <- sapply(data$ARREST_DATE, get_season)
data_2022 <- filter(data, year(ARREST_DATE) == 2022)

season_gender_counts22 <- data_2022 %>%
  group_by(Season, PERP_SEX) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  mutate(Season = factor(Season, levels = c("Spring", "Summer", "Autumn", "Winter")))

plotgender_22<- ggplot(season_gender_counts22, aes(x = Season, y = Count, fill = PERP_SEX)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("F" = "#f28e9b", "M" = "#8eaaf2")) +
  theme_minimal() +
  labs(title = "2022", 
       x = "Season", 
       y = "Number of Incidents",
       fill = "Gender")
```

```{r echo=F}
#2023 female vs male 
data$Season <- sapply(data$ARREST_DATE, get_season)
data_2023 <- filter(data, year(ARREST_DATE) == 2023)

season_gender_counts23 <- data_2023 %>%
  group_by(Season, PERP_SEX) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  mutate(Season = factor(Season, levels = c("Spring", "Summer", "Autumn", "Winter")))

other_genders <- unique(season_gender_counts23$PERP_SEX)

ggplot(season_gender_counts23, aes(x = Season, y = Count, fill = PERP_SEX)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7)) +
  scale_fill_manual(values = c("F" = "#ff9999", "M" = "#9999ff", "U" = "#ffff66")) +
  theme_light() +
  labs(title = "2023",
       x = "Season",
       y = "Count",
       fill = "Gender")
```

### Polish our graphs and charts

Titles and Labels: We are going to craft clear, descriptive titles and axis labels that succinctly convey the meaning of the data, and use numbers to label each in sequence.

Legends: We need to ensure legends are clear and correctly positioned. Modify legend titles to be informative and concise. Use a color scale that is intuitive and provides good contrast for the data represented.

Captions: we need to include captions that explain the visualization in a bit more detail, highlighting key findings or noting any caveats about the data.

Scales and Breaks: We decide to adjust scales and breaks on axes to make sure they are appropriate for the data’s range and distribution. 

Color Choices: We will ensure consistency across multiple visualizations, with the same color schemes, label formats, and styles used throughout.


We are planning to add a interactive graph that could brief will the crime variables in our dataset.



