---
title: "Fifth Blog Post"
subtitle: ""
author: "TEAM 14"
date: "2024-04-15"
draft: false
---

```{r include=F}
library(dplyr)
library(tidyverse) 
library(ggplot2)
library(dplyr)
library(kableExtra)
library(ggcorrplot)
library(ggmap)
library(leaflet)
library(magrittr)
library(sf)
ds <- read_rds("dataset/load_and_clean_data.rds") 
```

In our research, in addition to the explantory analysis of arrest data, we wanted to take it a step further and look at the non-human factors that affect crime, so we added the daily weather conditions of New York City over the course of a year to our master data. This data comes from [Visual crossing](https://www.visualcrossing.com/weather/weather-data-services#)'s weather database.


we selected basic weather conditions for New York City for the year 2023 so that we could explore whether extremes in temperature, wind speed, and so on, affect crime rates, as well as crime trends under different weather conditions. This way it can be more efficient to optimize the efficiency of NYPD's police response. 

[Data Loading and Cleaning Script](scripts/load_and_clean_data.R)


```{r echo =F}
crime_counts_by_day <- ds %>%
  mutate(ARREST_DATE = mdy(ARREST_DATE)) %>%  
  count(ARREST_DATE)  
  
crime_counts_by_month <- crime_counts_by_day %>%
  mutate(Month = floor_date(ARREST_DATE, "month")) %>%
  group_by(Month) %>%
  summarize(Monthly_Crime_Count = sum(n), .groups = "drop")

ggplot(crime_counts_by_month, aes(x = Month, y = Monthly_Crime_Count)) +
  geom_col(fill = "steelblue") +  # Use geom_col for bar charts
  labs(title = "Monthly Crime Counts 2023",
       x = "Month",
       y = "Crime Incidents") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) 
```
By examining the combined dataset through graphical analysis, we observed that there has been a steady increase in crime rates from winter to spring and decreases from Autumn to winter The proportion of different types of crimes remained relatively stable over the two years.

And our next step is to do the regression and logistic analysis to understanding the trends and patterns within the data. Moreover, we will create a more localized heat map of New York neighborhoods which will provide clear visual insights into areas with higher crime rates, potentially informing policy decisions, law enforcement resource allocation, or community support initiatives.









