---
title: "Fifth Blog Post"
subtitle: ""
author: "TEAM 14"
date: "2024-04-08"
draft: false
---

```{r include=F}
library(dplyr)
library(tidyverse) 
library(ggplot2)
library(dplyr)
library(kableExtra)
library(ggcorrplot)
library(ggmap)
library(leaflet)
library(magrittr)

ds <- read_rds("dataset/load_and_clean_data.rds") 
```

We combined NYPD Arrest Data from 2022 to our dataset. The new dataset of NYPD Arrest Data 2022 is similar to the previous dataset and they both include information about the type of crime, the location and time of enforcement. Initially, we ensured that both datasets have the same structure by matching the variables. We delete the arrest ID, two classification ID, and jurisdiction code. We remove rows with any NA and null values from the dataset. These variables are redundant and not relevant with our research. We also delete undefined data in Level of offense: felony, misdemeanor, violation, infraction.
Except that,After matching the variables and cleaning, the datasets were merged into a single dataset. 

Bellow is code of cleaning and merging process.

[Data Loading and Cleaning Script](scripts/load_and_clean_data.R)


```{r echo =F}
crime_counts_by_day <- ds %>%
  mutate(ARREST_DATE = mdy(ARREST_DATE)) %>%  
  count(ARREST_DATE)  
  
crime_counts_by_month <- crime_counts_by_day %>%
  mutate(Month = floor_date(ARREST_DATE, "month")) %>%
  group_by(Month) %>%
  summarize(Monthly_Crime_Count = sum(n), .groups = "drop")

ggplot(crime_counts_by_month, aes(x = Month, y = Monthly_Crime_Count)) +
  geom_col(fill = "steelblue") +  # Use geom_col for bar charts
  labs(title = "Monthly Crime Counts in 2022 to 2023",
       x = "Month",
       y = "Crime Incidents") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) 
```


By examining the combined dataset through graphical analysis, we observed that there has been a steady increase in crime rates from 2022 to 2023. The proportion of different types of crimes remained relatively stable over the two years.

And our next step is to do the regression and logistic analysis to understanding the trends and patterns within the data. Moreover, we will create a more localized heat map of New York neighborhoods which will provide clear visual insights into areas with higher crime rates, potentially informing policy decisions, law enforcement resource allocation, or community support initiatives.
